---
title: "STAT 892-004 Integrative Data Science for Plant Phenomics"
output: 
  learnr::tutorial
runtime: shiny_prerendered
---
```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


# Introduction to R 
This is an interacitive tutorial for [R](https://www.r-project.org/) based on the [learnr](https://cran.rstudio.com/web/packages/learnr/index.html) package. 
 
## Working directory
The function `getwd` returns your current working directory. This function has no argument. 

### Exercise 

Write the R code and find out your current directory. 

```{r ex1a, exercise=TRUE}

```

<div id="ex1a-hint">
**Hint:** Type `getwd()` and then click the blue button.
</div>





You can change the working directory by using `setwd` and specify the path to the desired directory. Make sure to put the directory name in double quotation marks `" "`.

### Exercise

Write the R code and set the working directory to your desired directory. Confirm that your working directory is set to your desired directory. 

```{r ex1b, exercise=TRUE}

```

```{r ex1b-solution}
setwd("~/Dropbox/chikudaisei/teaching/2018/STAT892/day03/")
getwd()
```

Note that it is likely that your path may differ from the path shown in the solution.





## File input

The function `read.table` reads a file in a data frame format. 

The R code below creates a file which contains two rows. The read.table function has many arguments but we will use only three in this exercise. The code then reads the file by using the `read.table` function and store the content in the variable `tmp`. 

```{r ex2a, exercise=TRUE}
cat("ID Pheno \n 1 3.7 \n 2 1.9 \n", file="tmp.txt", sep=" ")
tmp <- read.table(file = "tmp.txt", header = TRUE, stringsAsFactors = FALSE)
head(tmp)
```

The `head` function returns the first six rows of a data frame. The number of rows returned can be varied by controlling the argument `n`.

### Exercise

The function `read.table` can also read a file directly from WWW. Write the R code that reads the [simulated phenotype file](https://raw.githubusercontent.com/morota/asci431-2017/gh-pages/hw05/pheno.txt) and show the first 10 rows of the data. 
```{r ex2b, exercise=TRUE}

```


```{r ex2b-solution}
tmp <- read.table(file="https://raw.githubusercontent.com/morota/asci431-2017/gh-pages/hw05/pheno.txt", header = TRUE, stringsAsFactors = FALSE)
head(tmp, n=10)
```

Note that consider using the [readr](https://cran.r-project.org/web/packages/readr/index.html) or [data.table](https://cran.r-project.org/web/packages/data.table/index.html) to efficiently read large text files.






## Data structure 

Base R has a diverse data types. We will discuss 1) atomic vector, 2) list, 3) matrix, and 4) data frame in this exercize. 

### Atomic vector

Commonly used atomic vectors include integer, double (numeric), character, and logical. 

#### Integer 
```{r ex3a, exercise=TRUE}
x1 <- c(1L, 3L, 5L, 7L)
x1
class(x1)
```


#### Double (Numeric)
```{r ex3b, exercise=TRUE}
x2 <- c(1.5, 3.5, 5.5, 7.5)
x2
class(x2)
```


#### Character 
```{r ex3c, exercise=TRUE}
x3 <- c("Welcome", "to", "STAT 892-004", "Integrative Data Science for Plant Phenomics", "!")
x3
class(x3)
```

#### Logical 
```{r ex3d, exercise=TRUE}
x4 <- c(TRUE, FALSE)
x4
class(x4)
```

#### Vector index
We can access a certain element of a variable by specifying an index inside a single square bracket `[ ]` operator.

```{r ex3e, exercise=TRUE}
x3 <- c("Welcome", "to", "STAT 892-004", "Integrative Data Science for Plant Phenomics", "!")
x3[3]
```

### Exercise

Create an atomic vector variable containing both integer and double. Check the class of the variable. Interpret the result. 

```{r ex3f, exercise=TRUE}

```

```{r ex3f-solution}
x <- c(10L, 20.1)
class(x)
```

### Exercise

Create an atomic vector variable containing integer, double, and character. Check the class of the variable. Interpret the result. 

```{r ex3g, exercise=TRUE}

```

```{r ex3g-solution}
x <- c(10L, 20.1, "STAT892-004")
class(x)
```

### List

The advantage of lists is that it can contain any data types. Use the double square bracket  `[[ ]]` operator to access a list member. 
```{r ex3h, exercise=TRUE}
x5 <-list(10L, 20.1, "STAT892-004")
x5
class(x5)
x5[[2]]
```


### Exercise

```{r ex3i}
question("What is the class of variable x5 <- list(c(1L, 5L), c(1.5, 5.5), c(\"STAT 892-004\"), c(TRUE, FALSE))?",
    answer("logical"),
    answer("numeric"),
    answer("list", correct = TRUE),
    answer("character"),
     answer("integer"),
    allow_retry = TRUE
)
```


### Matrix

A matrix is a multi-dimensional vector or array. The function `matrix` creates a matrix variable. The `dim` function returns the dimension of a matrix. 

```{r ex3j, exercise=TRUE}
x6 <- matrix(c(1, 2, 3, 4), ncol=2, nrow=2)
x6
dim(x6)
class(x6)
x6[,1]
```


### Exercise
```{r ex3k}
question("What is the correct R syntax to access the 2nd row and 1st column of matrix?",
    answer("x6[1,1]"),
    answer("x6[1,2]"),
    answer("x6[2,1]", correct = TRUE),
    answer("x6[2,2]"),
    allow_retry = TRUE
)
```

Because a matrix is simply a multi-dimensional vector, it can not contain both integer/numeric and chracter. 

### Data frame

A data frame can be viewed as a mult-dimensional list of equal length vectors. This makes a data frame to include various data types in a single variable. 

```{r ex3l, exercise=TRUE}
x7 <- data.frame(int=c(1, 2, 3, 4), chr=c("A", "B", "C", "D"), stringsAsFactors = FALSE)
x7
```












## Statistical analysis

We will learn how to compute basic statistics in R by using the `PlantGrowth` data set. We can learn more about `PlantGrowth` by typing `?PlantGrowth` or `help(PlantGrowth)`.

### Exercise

Open the help page for `PlantGrowth` data set. 

```{r ex4a, exercise=TRUE}

```

```{r ex4a-solution}
?PlantGrowth
```

### Data structure 

This is what the `PlantGrowth` data set looks like.  
```{r ex4b, exercise=TRUE, exercise.eval = TRUE}
PlantGrowth
```

### Mean
The `mean` function computes the mean of a vector.
```{r ex4c, exercise=TRUE}
mean(PlantGrowth[,1])
```

### Variance
The `var` function computes the sample variance of a vector.
```{r ex4d, exercise=TRUE}
var(PlantGrowth[,1])
```

### Table
The `table` function computes the frequency of each of the discrete values.
```{r ex4e, exercise=TRUE}
table(PlantGrowth[,2])
```

### Summary
The `summary` function computes the summary of each column.
```{r ex4f, exercise=TRUE}
summary(PlantGrowth)
```

### Subsetting
Use the `subset` function to select a subset of data. 
```{r ex4g, exercise=TRUE}
subset(PlantGrowth, group == "ctrl")
```

### Exercise
Compute the mean of yields under `trt1` condition. 
```{r ex4h, exercise=TRUE}

```

```{r ex4h-solution}
mean(subset(PlantGrowth, group == "trt1")[,1])
```


### Exercise
Find out how many plants produced yields greater than 5, regardless of group type. 
```{r ex4i, exercise=TRUE}

```

```{r ex4i-solution}
dim(subset(PlantGrowth, weight > 5))
```


### Covariance
The `cov` function computes the covariance of two vectors. Covariance between yields under two different treatment conditions is as follows: 
```{r ex4j, exercise=TRUE}
cov(subset(PlantGrowth, group == "trt1")[,1], subset(PlantGrowth, group == "trt2")[,1]) 
```


### Correlation
The `cor` function computes the correlation of two vectors. Correlation between yields under two different treatment conditions is as follows: 
```{r ex4m, exercise=TRUE}
cor(subset(PlantGrowth, group == "trt1")[,1], subset(PlantGrowth, group == "trt2")[,1]) 
```


### Exercise
What is the correlation between yields of `ctrl` and `trt1`? 
```{r ex4n, exercise=TRUE}

```

```{r ex4n-solution}
cor(subset(PlantGrowth, group == "ctrl")[,1], subset(PlantGrowth, group == "trt1")[,1]) 
```

### Example function
The `example` function runs an example code in help pages. 
```{r ex4o, exercise=TRUE}
example("PlantGrowth")
```

